---
title: "Bioinformatics Class 7"
output: 
  html_document: 
    keep_md: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#functions again
we can source any file or F code with the `source`() function. 
```{r}
source("http://tinyurl.com/rescale-R")

```
lets make sure things are here

```{r}
ls()
```

check `rescale()` function is working

```{r}
rescale(1:10)
```


break 
```{r,eval=FALSE}
rescale( c(1:10,"string"))
```

Lets check if `rescale2()` does better 

```{r,eval=FALSE}
rescale2( c(1:10,"string"))
```
## Function for finding missing values in two datasets

write a `both_na()` function to do this 
```{r}
x <-  c(1, 2, NA, 3, NA)
y <-  c(NA, 3, NA, 3, 4)
is.na(x)

```
```{r}
which( is.na(x) )

```
```{r}
sum( is.na(x) )
```

using `&` operator and `sum()` to find how many things are missing in both experiments
```{r}
is.na(x) & is.na(y)
sum(is.na(x) & is.na(y))
```

now make function from this snippet. 

```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y) ) 
}
```

"eejit proof" testing 
```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

#both_na(x,y2)
#gives error!
```
using `both_na2()` with != operator

```{r, eval=FALSE}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
both_na2(x,y2)
```



