---
title: "lecture 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

generating example data for clustering
```{r}
tmp <-  c(rnorm(30,-3), rnorm(30,3))
x <- cbind( x=tmp, y=rev( tmp))

plot(x)
```

using kmeans center with k=2 and nstart=20
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```
 inspect/print results
 
```{r}
#cluster sie
km$size
#cluster assignment
km$cluster
#cluster centers
km$centers
```
 plot x colored by kmeans assignment ad add cluster centers
 
```{r}
plot(x, col=km$cluster, pch=15)
points(km$centers, col="blue", pch=16)
```

Hierachical clustering
```{r}
# First we need to calculate point (dis)similarity
# as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
# clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc 
```
 
regular viewing will not work, `dist()` does not output into data frame. 

```{r}
View( as.matrix(dist_matrix))
dim(as.matrix(dist_matrix))
```

plotting
```{r}
plot(hc)
```

adding abline
and cutting tree to define clusters
```{r}
plot(hc)
abline(h=6, col="red")
#can either cut by h or by k 
grps <- cutree(hc, h=6)
table( grps)
plot(x, col=grps) # returns same colors as k means clustering above!
```


try different cutting

```{r}
plot(x, col=cutree(hc, k=3))
```

```{r}
# Draws a dendrogram
plot(hc)
abline(h=6, col="red")
grps2<- cutree(hc, k=2 ) # Cut into k grps
plot(hc, col=grps2)
```

Trying hierarchal clustering

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
 matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
 matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
 matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
 rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x)
# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)

```

my attempts
```{r}
my.dist.matrix<- dist(x)
hc2<- hclust(my.dist.matrix)
grps2<- cutree(hc2, k = 3)
plot(hc2)
table(grps2)

```

plotting data colored by cluster

```{r}
plot(x, col=grps2)
```

